---
title: "Relational Data"
author: "Ryan Ross"
date: "October 31, 2017"
output: html_document
---

```{r}
library(tidyverse)
library(nycflights13)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 13.2.1 Exercises


1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

Variables: flights, airports
tables: origin, dest
Merge flights with airports twice to find out the location of the origin and destination airports.

2. I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

The variable origin in weather is matched with faa in airports.

3. weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

year, month, day, hour, origin in weather would be matched to year, month, day, hour, dest in flight

4. We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

I would create a table for special dates. The primary key would be date. It would match to the year, month, day columns of flights.

# 13.3.1 Exercises

1. Add a surrogate key to flights.


```{r}
flights %>% 
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) %>%
  glimpse()
```


Identify the keys in the following datasets

Lahman::Batting - playerID, yearID, stint
babynames::babynames - year, sex, name
nasaweather::atmos - lat, long, year, month
fueleconomy::vehicles - id
ggplot2::diamonds - N/A

2. Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers. ????

3. How would you characterise the relationship between the Batting, Pitching, and Fielding tables?


# 13.4.6 Exercises

1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.


```{r}
destinationDelays <-
  flights %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa"))

destinationDelays %>%
  ggplot(aes(lon, lat, color = delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```


2. Add the location of the origin and destination (i.e. the lat and lon) to flights.

```{r}
flights %>%
  left_join(airports, by = c(dest = "faa")) %>%
  left_join(airports, by = c(origin = "faa")) %>%
  head()
```


3. Is there a relationship between the age of a plane and its delays?

```{r}
plane_ages <- 
  planes %>%
  mutate(age = 2013 - year) %>%
  select(tailnum, age)

flights %>%
  inner_join(plane_ages, by = "tailnum") %>%
  group_by(age) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean(dep_delay)) %>%
  ggplot(aes(x = age, y = delay)) +
  geom_point() +
  geom_line()
```


There is no relationship between age and delay. However, I believe that could be due to the plane being used less frequently. 

4. What weather conditions make it more likely to see a delay?

```{r}
flight_weather <-
  flights %>%
  inner_join(weather, by = c("origin" = "origin","year" = "year", "month" = "month", "day" = "day", "hour" = "hour"))

flight_weather %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = precip, y = delay)) +
    geom_line() + geom_point()
```

Seems as if precipitation causes delays. There isn't a relationship for amount of precipitation though, it's pretty sporadic.

5. What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

```{r}
library(viridis)
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, color = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() + 
  scale_color_viridis()
```


# 13.5.1 Exercises

1. What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

```{r}
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(carrier, sort = TRUE)
```


2. Filter flights to only show flights with planes that have flown at least 100 flights.

```{r}
planes100 <- 
  filter(flights) %>%
  group_by(tailnum) %>%
  count() %>%
  filter(n > 100)

flights %>%
  semi_join(planes100, by = "tailnum")
```


3. Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

```{r}
glimpse(fueleconomy::vehicles)
```
```{r}
glimpse(fueleconomy::common)
```

4. Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

```{r}
planes48delay <- 
  filter(flights) %>%
  group_by(dep_delay) %>%
  count() %>%
  filter(dep_delay > 1000)

flights %>%
  semi_join(planes48delay, by = "dep_delay")
```


```{r}
planes48delay2 <- 
  filter(flight_weather) %>%
  group_by(dep_delay) %>%
  count() %>%
  filter(dep_delay > 1000)

flights %>%
  semi_join(planes48delay2, by = "dep_delay")
```

5. What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

anti_join(flights, airports, by = c("dest" = "faa")) are flights that go to an airport that is not in FAA list of destinations, likely foreign airports.

anti_join(airports, flights, by = c("faa" = "dest")) are US airports that don’t have a flight in the data, meaning that there were no flights to that aiport from New York in 2013.

6. You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

I reject, because there is no relationship between plane and airline.











